## Feature Selection with Forward Sequential Selection (FSS)

This example uses FSS (forward sequential selection) to build a subset of features by adding, at each step, the feature that most improves the evaluation criterion.

Prerequisites
- R packages: daltoolbox, daltoolboxdp

```{r}
# Feature Selection

# Installation (if needed)
#install.packages("daltoolboxdp")
```

```{r}
# Loading packages
library(daltoolbox)
library(daltoolboxdp)
```


```{r}
# Example data
iris <- datasets::iris
```

```{r}
# FSS - step by step

# 1) Fit the selector with target "Species"
myfeature <- fit(fs_fss("Species"), iris)

# 2) View selected features
print(myfeature$features)

# 3) Transform data to keep selected features + target
data <- transform(myfeature, iris)
print(head(data))
```

### Method

Forward Sequential Selection starts from an empty set and iteratively adds the feature that most improves a chosen criterion (e.g., validation accuracy), stopping by a rule or when performance no longer improves.

### References
- Whitney, A. W. (1971). A direct method of nonparametric measurement selection. IEEE Trans. Computers, 20(9), 1100–1103.

