## Class Balance: Subsampling

Random undersampling reduces the majority class by removing examples until the classes are balanced. It can mitigate bias toward the majority class but may discard useful information if too aggressive.

This example shows how to handle class imbalance by applying subsampling (reduce the majority class) on an imbalanced subset of the Iris dataset.

Prerequisites
- R packages: daltoolbox, daltoolboxdp

```{r}
# Installation (if needed)
#install.packages("daltoolboxdp")
```

```{r}
# Loading packages
library(daltoolbox)
library(daltoolboxdp)
```

```{r}
# Example data and creation of artificial imbalance
iris <- datasets::iris
data(iris)
mod_iris <- iris[c(1:50,51:71,101:111),]
table(mod_iris$Species)                      # original distribution
```

```{r}
# Subsampling - reduce the majority class to balance
bal <- bal_subsampling('Species')
bal <- daltoolbox::fit(bal, mod_iris)
adjust_iris <- daltoolbox::transform(bal, mod_iris)
table(adjust_iris$Species)                    # distribution after subsampling
```

References
- He, H., & Garcia, E. A. (2009). Learning from imbalanced data. IEEE Transactions on Knowledge and Data Engineering, 21(9), 1263–1284.

