## 08. PyTorch MLP Classifier with Dynamic Validation and Patience

This example focuses on the training regime rather than only the network architecture. It shows how to train the PyTorch-backed MLP classifier with a dynamic validation split and patience-based early stopping.

Prerequisites
- R packages: daltoolbox, daltoolboxdp
- Python with PyTorch accessible via reticulate

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/daltoolboxdp/main/examples/seed.R"))
# Installation (if needed)
#install.packages("daltoolboxdp")
```

```{r}
library(daltoolbox)
library(daltoolboxdp)
```

```{r}
# Loading Iris dataset
iris <- datasets::iris
```

```{r}
# Train/test split
slevels <- levels(iris$Species)

set.seed(1)
sr <- sample_random()
sr <- train_test(sr, iris)
iris_train <- sr$train
iris_test <- sr$test
```

```{r}
# Dynamic validation with patience-based early stopping
model <- torch_cla_mlp(
  attribute = "Species",
  slevels = slevels,
  input_size = 4L,
  hidden_sizes = c(16L, 8L),
  num_classes = 3L,
  epochs = 300L,
  validation_strategy = "dynamic",
  stopping_rule = "patience",
  patience = 20L,
  val_ratio = 0.2
)

set_example_seed()
model <- fit(model, iris_train)
```

Training configuration
- `validation_strategy = "dynamic"` redraws the train/validation split at each epoch.
- `stopping_rule = "patience"` stops training after a chosen number of epochs without meaningful validation improvement.
- `epochs = 300L` is an upper bound here; the effective number of epochs is controlled by early stopping.

```{r}
# Training evaluation
train_prediction <- predict(model, iris_train)
iris_train_predictand <- adjust_class_label(iris_train[, "Species"])
train_eval <- evaluate(model, iris_train_predictand, train_prediction)
print(train_eval$metrics)
```

```{r}
# Test evaluation
test_prediction <- predict(model, iris_test)
iris_test_predictand <- adjust_class_label(iris_test[, "Species"])
test_eval <- evaluate(model, iris_test_predictand, test_prediction)
print(test_eval$metrics)
```

```{r}
# Effective training duration
print(model$epochs_done)
```

```{r}
# Training and validation curves
fit_loss <- data.frame(
  x = seq_along(model$train_loss_hist),
  train_loss = model$train_loss_hist
)

if (!is.null(model$val_loss_hist) && length(model$val_loss_hist) > 0) {
  fit_loss$val_loss <- model$val_loss_hist
}

colors <- if ("val_loss" %in% names(fit_loss)) c("Blue", "Orange") else c("Blue")
grf <- plot_series(fit_loss, colors = colors)
plot(grf)
```

Notes
- This example is useful when a fixed validation partition feels too brittle on smaller datasets.
- To compare behaviors, keep the same architecture and swap only `stopping_rule` among `"patience"`, `"sma"`, `"ema"`, and `"h"`.
- To return to the default baseline, set `validation_strategy = "static"` and `stopping_rule = "none"`.

References
- Rumelhart, D. E., Hinton, G. E., & Williams, R. J. (1986). Learning representations by back-propagating errors.
- Paszke, A., et al. (2019). PyTorch: An Imperative Style, High-Performance Deep Learning Library.
