## PyTorch Multi-Layer Perceptron (MLP) Classifier

This example uses the PyTorch-backed MLP classifier exposed by `daltoolboxdp` to classify the Iris dataset. The workflow mirrors the scikit-learn MLP example: split train/test, train, predict, and evaluate.

Prerequisites
- R packages: daltoolbox, daltoolboxdp
- Python with PyTorch accessible via reticulate

```{r}
# Installation (if needed)
#install.packages("daltoolboxdp")
```

```{r}
library(daltoolbox)
library(daltoolboxdp)
```

```{r}
# Loading Iris dataset
iris <- datasets::iris
```

```{r}
# Training and evaluation with PyTorch MLP
slevels <- levels(iris$Species)

set.seed(1)
sr <- sample_random()
sr <- train_test(sr, iris)
iris_train <- sr$train
iris_test <- sr$test

model <- torch_cla_mlp(
  attribute = "Species",
  slevels = slevels,
  input_size = 4L,
  hidden_sizes = c(16L, 8L),
  num_classes = 3L,
  epochs = 100L
)

model <- fit(model, iris_train)
train_prediction <- predict(model, iris_train)

iris_train_predictand <- adjust_class_label(iris_train[, "Species"])
train_eval <- evaluate(model, iris_train_predictand, train_prediction)
print(train_eval$metrics)
```

```{r}
# Test prediction and evaluation
test_prediction <- predict(model, iris_test)

iris_test_predictand <- adjust_class_label(iris_test[, "Species"])
test_eval <- evaluate(model, iris_test_predictand, test_prediction)
print(test_eval$metrics)
```

```{r}
# Predicted probabilities
probabilities <- predict_proba.torch_cla_mlp(model, iris_test[, !names(iris_test) %in% "Species"])
head(probabilities)
```

Notes
- By default, this example uses `validation_strategy = "static"` and `stopping_rule = "none"`.
- To enable early stopping, change `stopping_rule` to `"patience"`, `"sma"`, `"ema"`, or `"h"`.
- To switch to the dynamic split, use `validation_strategy = "dynamic"`.

References
- Rumelhart, D. E., Hinton, G. E., & Williams, R. J. (1986). Learning representations by back-propagating errors.
- Paszke, A., et al. (2019). PyTorch: An Imperative Style, High-Performance Deep Learning Library.
