## LSTM Autoencoder (encode)

LSTM autoencoders use recurrent units to encode temporal dependencies in a sequence (window) into a fixed-sized latent vector. The memory cell and gating mechanisms help capture long- and short-term patterns before decoding or downstream use of the latent code.

This example demonstrates the use of an LSTM-based Autoencoder to encode windows of a time series. The LSTM encoder learns sequence representations, reducing from p to k dimensions.

Prerequisites
- Python with PyTorch accessible via reticulate
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2

Quick notes
- Architecture: LSTM encoder + MLP/decoder to represent time-series windows.
- Useful when there is strong temporal dependence within each window.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/daltoolboxdp/main/examples/seed.R"))
# Installing example dependencies (if needed)
#install.packages("tspredit")
#install.packages("daltoolboxdp")
```

```{r}
# Loading required packages
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
# Example dataset (series -> windows)
data(tsd)

sw_size <- 5                      # sliding window size (p)
ts <- ts_data(tsd$y, sw_size)     # convert series into windows with p columns

ts_head(ts)
```

```{r}
# Normalization (min-max by group)
preproc <- ts_norm_gminmax()
set_example_seed()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

ts_head(ts)
```

```{r}
# Train/test split
samp <- ts_sample(ts, test_size = 10)
train <- as.data.frame(samp$train)
test  <- as.data.frame(samp$test)
```

```{r}
# Creating the LSTM autoencoder: reduce from 5 -> 3 dimensions (p -> k)
# - this example intentionally increases the number of epochs above the library default
# - use the full window as a sequence and a larger recurrent hidden state
auto <- autoenc_lstm_e(
  5, 3,
  sequence_length = 5L,
  lstm_hidden_size = 16L,
  epochs = 200L
)

# Training the model
set_example_seed()
auto <- fit(auto, train)
```

Constructor configuration
- Fixed-epoch baseline: omit `epochs` to use the default value of `100L` and keep `validation_strategy = "static"` with `stopping_rule = "none"`.
- Static early stopping: keep `validation_strategy = "static"` and choose `stopping_rule = "patience"`, `"sma"`, `"ema"`, or `"h"`.
- Dynamic early stopping: switch `validation_strategy = "dynamic"` and reuse the same stopping rules.
- The loss plot below always shows `train_loss`; it adds `val_loss` when validation is active.

Architecture variations
- `sequence_length` reshapes each row into a longer temporal sequence before the recurrent encoder.
- `lstm_hidden_size` decouples recurrent capacity from the latent bottleneck `encoding_size`.
- `num_layers` and `dropout` enable deeper recurrent autoencoders.
- For this example, `sequence_length = 5L` makes each window a true 5-step sequence, which is usually more informative than the degenerate `sequence_length = 1L`.

```{r}
# Learning curves
fit_loss <- data.frame(
  x = seq_along(auto$train_loss),
  train_loss = auto$train_loss
)
if (!is.null(auto$val_loss) && length(auto$val_loss) > 0) {
  fit_loss$val_loss <- auto$val_loss
}
colors <- if ("val_loss" %in% names(fit_loss)) c("Blue", "Orange") else c("Blue")
grf <- plot_series(fit_loss, colors = colors)
plot(grf)
```

```{r}
# Testing the autoencoder (encoding)
# Show samples from the test set and the encoding (k columns)
print(head(test))
result <- transform(auto, test)
print(head(result))
```

References
- Hochreiter, S., & Schmidhuber, J. (1997). Long short-term memory. Neural Computation, 9(8), 1735–1780.

