## Variational Autoencoder (encode)

Variational Autoencoders learn a probabilistic encoder that outputs parameters of a latent distribution (e.g., mean and variance) and a decoder that reconstructs from latent samples. The loss combines reconstruction error and a KL divergence that regularizes the latent distribution toward a prior.

This example uses a Variational Autoencoder (VAE) to learn latent representations of time-series windows. The VAE reduces from p to k dimensions and regularizes the latent space to approximate a target distribution (e.g., standard normal) via a KL term.

Prerequisites
- Python with PyTorch accessible via reticulate
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2

Quick notes
- Loss: reconstruction + KL divergence between the latent distribution and the prior.
- Useful for generating continuous, well-behaved representations in latent space.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/daltoolboxdp/main/examples/seed.R"))
# Installing example dependencies (if needed)
#install.packages("tspredit")
#install.packages("daltoolboxdp")
```

```{r}
# Loading required packages
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
# Example dataset (series -> windows)
data(tsd)

sw_size <- 5                      # sliding window size (p)
ts <- ts_data(tsd$y, sw_size)     # convert series into windows with p columns

ts_head(ts)
```

```{r}
# Normalization (min-max by group)
preproc <- ts_norm_gminmax()
set_example_seed()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

ts_head(ts)
```

```{r}
# Train/test split
samp <- ts_sample(ts, test_size = 10)
train <- as.data.frame(samp$train)
test  <- as.data.frame(samp$test)
```

```{r}
# Creating the VAE: reduce from 5 -> 3 dimensions (p -> k)
# - the default number of epochs is used
auto <- autoenc_variational_e(5, 3)

# Training the model
set_example_seed()
auto <- fit(auto, train)
```

Constructor configuration
- Fixed-epoch baseline: omit `epochs` to use the default value and keep `validation_strategy = "static"` with `stopping_rule = "none"`.
- Static early stopping: keep `validation_strategy = "static"` and choose `stopping_rule = "patience"`, `"sma"`, `"ema"`, or `"h"`.
- Dynamic early stopping: switch `validation_strategy = "dynamic"` and reuse the same stopping rules.
- The loss plot below always shows `train_loss`; it adds `val_loss` when validation is active.

Architecture variations
- The original VAE is recovered with `encoder_hidden_sizes = c(64L, 32L)`.
- Deeper VAEs can be built with larger stacks such as `c(128L, 64L, 32L)`.
- `reconstruction_loss = "bce"` works well with normalized inputs and `output_activation = "sigmoid"`.
- `reconstruction_loss = "mse"` is often preferable for unconstrained real-valued windows.

```{r}
# Learning curves
fit_loss <- data.frame(
  x = seq_along(auto$train_loss),
  train_loss = auto$train_loss
)
if (!is.null(auto$val_loss) && length(auto$val_loss) > 0) {
  fit_loss$val_loss <- auto$val_loss
}
colors <- if ("val_loss" %in% names(fit_loss)) c("Blue", "Orange") else c("Blue")
grf <- plot_series(fit_loss, colors = colors)
plot(grf)
```

```{r}
# Testing the VAE (encoding)
# Show samples from the test set and the encoding (k columns)
print(head(test))
result <- transform(auto, test)
print(head(result))
```

References
- Kingma, D. P., & Welling, M. (2014). Auto-Encoding Variational Bayes. ICLR.

