## Denoising Autoencoder (encode-decode)

Inputs are stochastically corrupted during training, and the model learns to reconstruct clean windows. The bottleneck captures noise-invariant structure, and reconstruction metrics quantify how well essential patterns are preserved.

This example demonstrates how to use a denoising autoencoder to encode and reconstruct time-series windows, enabling evaluation of reconstruction quality under noise.

Prerequisites
- Python with PyTorch accessible via reticulate
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2

Quick notes
- Noise is applied to the input during training; at inference, the reconstruction tends to be smoother.
- Per-column metrics (R2 and MAPE) help assess robustness per step.

```{r}
source(url("https://raw.githubusercontent.com/cefet-rj-dal/daltoolboxdp/main/examples/seed.R"))
# Denoising Autoencoder transformation (encode-decode)

# Considering a dataset with $p$ numerical attributes. 

# The goal of the autoencoder is to reduce the dimension of $p$ to $k$, such that these $k$ attributes are enough to recompose the original $p$ attributes. However, from the $k$ dimensions the data is returned back to $p$ dimensions. The higher the autoencoder quality, the more similar the output is to the input. 

# Installing packages
#install.packages("tspredit")
#install.packages("daltoolboxdp")
```

```{r}
# Loading packages
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
# Example dataset (series -> windows) 
data(tsd)

sw_size <- 5
ts <- ts_data(tsd$y, sw_size)

ts_head(ts)
```

```{r}
# Normalization (min-max by group)
preproc <- ts_norm_gminmax()
set_example_seed()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

ts_head(ts)
```

```{r}
# Train/test split
samp <- ts_sample(ts, test_size = 10)
train <- as.data.frame(samp$train)
test <- as.data.frame(samp$test)
```

```{r}
# Training autoencoder (reduce 5 -> 3)
auto <- autoenc_denoise_ed(5, 3)
set_example_seed()
auto <- fit(auto, train)
```

Constructor configuration
- Fixed-epoch baseline: omit `epochs` to use the default value and keep `validation_strategy = "static"` with `stopping_rule = "none"`.
- Static early stopping: keep `validation_strategy = "static"` and choose `stopping_rule = "patience"`, `"sma"`, `"ema"`, or `"h"`.
- Dynamic early stopping: switch `validation_strategy = "dynamic"` and reuse the same stopping rules.
- The loss plot below always shows `train_loss`; it adds `val_loss` when validation is active.

Architecture variations
- Use the same dense controls as `autoenc_denoise_e`, plus `output_activation` when reconstruction scale should be constrained.

```{r}
fit_loss <- data.frame(x = seq_along(auto$train_loss), train_loss = auto$train_loss)
if (!is.null(auto$val_loss) && length(auto$val_loss) > 0) {
  fit_loss$val_loss <- auto$val_loss
}

colors <- if ("val_loss" %in% names(fit_loss)) c("Blue", "Orange") else c("Blue")
grf <- plot_series(fit_loss, colors = colors)
plot(grf)
```

```{r}
# Testing the autoencoder
# Show test samples and display reconstruction
print(head(test))
result <- transform(auto, test)
print(head(result))
```

```{r}
# Reconstruction metrics per column: R2 and MAPE
result <- as.data.frame(result)
names(result) <- names(test)
r2 <- c()
mape <- c()
for (col in names(test)){
  r2_col <- cor(test[col], result[col])^2
  r2 <- append(r2, r2_col)
  mape_col <- mean((abs((result[col] - test[col]))/test[col])[[col]])
  mape <- append(mape, mape_col)
  print(paste(col, 'R2 test:', r2_col, 'MAPE:', mape_col))
}
print(paste('Means R2 test:', mean(r2), 'MAPE:', mean(mape)))
```
 
```{r}
# Note: the noise level impacts reconstruction capacity.
```

References
- Vincent, P., Larochelle, H., Bengio, Y., & Manzagol, P. A. (2008). Extracting and composing robust features with denoising autoencoders. ICML.

