## 21. Autoencoder (Encode-Decode) with Static Validation and Patience

This example keeps the simple dense autoencoder architecture, but changes the training regime to use a fixed validation split and patience-based early stopping. It is useful when reconstruction quality matters and you want a stable reference validation subset across epochs.

Prerequisites
- Reticulate configured and Python with PyTorch installed
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2

```{r}
# Installing packages
#install.packages("tspredit")
#install.packages("daltoolboxdp")
```

```{r}
# Loading packages
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
# Example dataset (series -> windows)
data(tsd)

sw_size <- 5
ts <- ts_data(tsd$y, sw_size)

preproc <- ts_norm_gminmax()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

samp <- ts_sample(ts, test_size = 10)
train <- as.data.frame(samp$train)
test <- as.data.frame(samp$test)
```

```{r}
# Static validation with patience-based early stopping
auto <- autoenc_ed(
  5, 3,
  epochs = 300L,
  validation_strategy = "static",
  stopping_rule = "patience",
  patience = 20L,
  val_ratio = 0.2
)
auto <- fit(auto, train)
```

Training configuration
- `validation_strategy = "static"` keeps the same validation partition during the whole training run.
- `stopping_rule = "patience"` stops training when validation loss does not improve for a chosen number of epochs.
- `epochs = 300L` is only an upper bound; the effective number of epochs is determined by early stopping.

```{r}
# Effective training duration
print(length(auto$train_loss))
```

```{r}
# Training and validation curves
fit_loss <- data.frame(x = seq_along(auto$train_loss), train_loss = auto$train_loss)
if (!is.null(auto$val_loss) && length(auto$val_loss) > 0) {
  fit_loss$val_loss <- auto$val_loss
}

colors <- if ("val_loss" %in% names(fit_loss)) c("Blue", "Orange") else c("Blue")
grf <- plot_series(fit_loss, colors = colors)
plot(grf)
```

```{r}
# Testing: reconstruction of the test set
print(head(test))
result <- transform(auto, test)
result <- as.data.frame(result)
names(result) <- names(test)
print(head(result))
```

```{r}
# Evaluating reconstruction quality: R2 and MAPE per attribute
r2 <- c()
mape <- c()
for (col in names(test)){
  r2_col <- cor(test[col], result[col])^2
  r2 <- append(r2, r2_col)
  mape_col <- mean((abs((result[col] - test[col]))/test[col])[[col]])
  mape <- append(mape, mape_col)
  print(paste(col, 'R2 test:', r2_col, 'MAPE:', mape_col))
}
print(paste('Means R2 test:', mean(r2), 'MAPE:', mean(mape)))
```

Notes
- This setup is easier to interpret because all epochs are judged against the same validation subset.
- To compare with redrawn validation splits, see `22_autoenc_ed_dynamic_patience`.

References
- Goodfellow, I., Bengio, Y., & Courville, A. (2016). Deep Learning. MIT Press. (Chapter on Autoencoders)
