## Denoising Autoencoder (encode)

This example demonstrates how to use a denoising autoencoder to learn a robust encoding of time-series windows. During training, noise is added to the input.

Prerequisites
- Python with PyTorch accessible via reticulate
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2

Quick notes
- Idea: add noise to the input and train the model to recover the clean signal.
- Benefit: latent representations that are more robust to perturbations.

```{r}
# Denoising Autoencoder transformation (encode)

# Considering a dataset with $p$ numerical attributes. 

# The goal of the autoencoder is to reduce the dimension of $p$ to $k$, such that these $k$ attributes are enough to recompose the original $p$ attributes. 

# Installing packages
#install.packages("tspredit")
#install.packages("daltoolboxdp")
```

```{r}
# Loading packages
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
# Example dataset (series -> windows) 
data(tsd)

sw_size <- 5
ts <- ts_data(tsd$y, sw_size)

ts_head(ts)
```

```{r}
# Normalization (min-max by group)
preproc <- ts_norm_gminmax()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

ts_head(ts)
```

```{r}
# Train/test split
samp <- ts_sample(ts, test_size = 10)
train <- as.data.frame(samp$train)
test <- as.data.frame(samp$test)
```

```{r}
# Training autoencoder (reduce 5 -> 3)
auto <- autoenc_denoise_e(5, 3, num_epochs=1500)
auto <- fit(auto, train)
```

```{r}
fit_loss <- data.frame(x=1:length(auto$train_loss), train_loss=auto$train_loss,val_loss=auto$val_loss)

grf <- plot_series(fit_loss, colors=c('Blue','Orange'))
plot(grf)
```
 
```{r}
# Convergence should be stable; too much noise can hinder fitting.
```

```{r}
# Testing the autoencoder
# Show test samples and display encoding
print(head(test))
result <- transform(auto, test)
print(head(result))
```

