## Stacked Autoencoder (encode)

Stacked autoencoders compose multiple nonlinear layers to gradually compress inputs into a low-dimensional code. Depth allows the model to learn hierarchical features before exposing the compact representation for downstream use.

This example demonstrates a Stacked Autoencoder for encoding time-series windows, reducing from p to k dimensions with successive dense layers.

Prerequisites
- Python with PyTorch accessible via reticulate
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2

```{r}
# Installing example dependencies (if needed)
#install.packages("tspredit")
#install.packages("daltoolboxdp")
```

```{r}
# Loading required packages
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
# Example dataset (series -> windows)
data(tsd)

sw_size <- 5                      # sliding window size (p)
ts <- ts_data(tsd$y, sw_size)     # convert series into windows with p columns

ts_head(ts)
```

```{r}
# Normalization (min-max by group)
preproc <- ts_norm_gminmax()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

ts_head(ts)
```

```{r}
# Train/test split
samp <- ts_sample(ts, test_size = 10)
train <- as.data.frame(samp$train)
test  <- as.data.frame(samp$test)
```

```{r}
# Creating the stacked autoencoder: reduce from 5 -> 3 dimensions (p -> k)
auto <- autoenc_stacked_e(5, 3)

# Training the model
auto <- fit(auto, train)
```

```{r}
# Learning curves
fit_loss <- data.frame(
  x = seq_along(auto$train_loss),
  train_loss = auto$train_loss
)
if (!is.null(auto$val_loss) && length(auto$val_loss) > 0) {
  fit_loss$val_loss <- auto$val_loss
}
colors <- if ("val_loss" %in% names(fit_loss)) c("Blue", "Orange") else c("Blue")
grf <- plot_series(fit_loss, colors = colors)
plot(grf)
```

```{r}
# Testing the autoencoder (encoding)
# Show samples from the test set and the encoding (k columns)
print(head(test))
result <- transform(auto, test)
print(head(result))
```

References
- Bengio, Y., Lamblin, P., Popovici, D., & Larochelle, H. (2007). Greedy layer-wise training of deep networks. NIPS.
