## Variational Autoencoder (encode)

This example uses a Variational Autoencoder (VAE) to learn latent representations of time-series windows. The VAE reduces from p to k dimensions and regularizes the latent space to approximate a target distribution (e.g., standard normal) via a KL term.

Prerequisites
- Python with PyTorch accessible via reticulate
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2

Quick notes
- Loss: reconstruction + KL divergence between the latent distribution and the prior.
- Useful for generating continuous, well-behaved representations in latent space.

```{r}
# Installing example dependencies (if needed)
#install.packages("tspredit")
#install.packages("daltoolboxdp")
```

```{r}
# Loading required packages
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
# Example dataset (series -> windows)
data(tsd)

sw_size <- 5                      # sliding window size (p)
ts <- ts_data(tsd$y, sw_size)     # convert series into windows with p columns

ts_head(ts)
```

```{r}
# Normalization (min-max by group)
preproc <- ts_norm_gminmax()
preproc <- fit(preproc, ts)
ts <- transform(preproc, ts)

ts_head(ts)
```

```{r}
# Train/test split
samp <- ts_sample(ts, test_size = 10)
train <- as.data.frame(samp$train)
test  <- as.data.frame(samp$test)
```

```{r}
# Creating the VAE: reduce from 5 -> 3 dimensions (p -> k)
# - num_epochs: fewer epochs may suffice given the additional KL term
auto <- autoenc_variational_e(5, 3, num_epochs = 350)

# Training the model
auto <- fit(auto, train)
```

```{r}
# Learning curves (total loss per epoch)
fit_loss <- data.frame(
  x = 1:length(auto$train_loss),
  train_loss = auto$train_loss,
  val_loss = auto$val_loss
)
grf <- plot_series(fit_loss, colors = c('Blue', 'Orange'))
plot(grf)
```

```{r}
# Testing the VAE (encoding)
# Show samples from the test set and the encoding (k columns)
print(head(test))
result <- transform(auto, test)
print(head(result))
```

