## Time Series Forecasting with LSTM

This example shows how to use the PyTorch-backed `ts_lstm` forecaster. The workflow follows the `tspredit` sliding-window approach: transform a univariate series into lagged windows, fit the model, inspect training curves, and generate one-step-ahead predictions on the hold-out windows.

Prerequisites
- R packages: daltoolbox, tspredit, daltoolboxdp, ggplot2
- Python with PyTorch accessible via reticulate

```{r}
library(daltoolbox)
library(tspredit)
library(daltoolboxdp)
library(ggplot2)
```

```{r}
data(tsd)

sw_size <- 5
ts <- data.frame(unclass(ts_data(tsd$y, sw_size)))

split_at <- nrow(ts) - 5
train <- ts[1:split_at, ]
test <- ts[(split_at + 1):nrow(ts), ]
```

```{r}
model <- ts_lstm(
  preprocess = ts_norm_gminmax(),
  input_size = 4L,
  epochs = 100L
)

model <- fit(model, train[, 1:4], train[, 5])
```

```{r}
# Training curves
fit_loss <- data.frame(
  x = seq_along(model$train_loss_hist),
  train_loss = model$train_loss_hist
)
if (!is.null(model$val_loss_hist) && length(model$val_loss_hist) > 0) {
  fit_loss$val_loss <- model$val_loss_hist
}

colors <- if ("val_loss" %in% names(fit_loss)) c("Blue", "Orange") else c("Blue")
grf <- plot_series(fit_loss, colors = colors)
plot(grf)
```

```{r}
# One-step-ahead prediction on the test windows
prediction <- predict(model, test[, 1:4])
print(prediction)
```

Notes
- The default configuration is `validation_strategy = "static"` and `stopping_rule = "none"`, so only the training curve is shown.
- To display validation loss as well, use an early-stopping rule such as `"patience"`, `"sma"`, `"ema"`, or `"h"`.

References
- Hochreiter, S., & Schmidhuber, J. (1997). Long short-term memory.
