## PyTorch MLP Regressor

This example shows how to train a PyTorch-backed MLP regressor with `daltoolboxdp`. The example uses `mtcars` and follows a standard regression workflow: split, fit, predict, and evaluate.

Prerequisites
- R packages: daltoolbox, daltoolboxdp
- Python with PyTorch accessible via reticulate

```{r}
# Installation (if needed)
#install.packages("daltoolboxdp")
```

```{r}
library(daltoolbox)
library(daltoolboxdp)
```

```{r}
data <- mtcars

set.seed(1)
idx <- sample(seq_len(nrow(data)), size = floor(0.8 * nrow(data)))
train <- data[idx, ]
test <- data[-idx, ]
```

```{r}
# Fit the regressor
model <- torch_reg_mlp(
  attribute = "mpg",
  input_size = ncol(data) - 1L,
  hidden_sizes = c(16L, 8L),
  epochs = 100L
)

model <- fit(model, train)
```

```{r}
# Test prediction
prediction <- predict(model, test)
head(prediction)
```

```{r}
# Regression evaluation
test_eval <- evaluate(model, test$mpg, prediction)
print(test_eval$metrics)
```

Notes
- Default configuration uses `validation_strategy = "static"` and `stopping_rule = "none"`.
- To activate early stopping, set `stopping_rule` to `"patience"`, `"sma"`, `"ema"`, or `"h"`.
- To activate dynamic validation splits, use `validation_strategy = "dynamic"`.

References
- Bishop, C. M. (1995). Neural Networks for Pattern Recognition.
- Paszke, A., et al. (2019). PyTorch: An Imperative Style, High-Performance Deep Learning Library.
