About the method
- `torch_reg_mlp`: PyTorch-backed multilayer perceptron for regression.
- Hyperparameters: `hidden_sizes` (hidden-layer width/depth), `dropout`, `epochs`, `lr`, and the unified validation/early-stopping controls.

Didactic goal: read this example as the same numeric-prediction workflow used by `reg_mlp`. The Experiment Line stays the same; the practical change is the constructor.

Environment setup.
```{r}
# Regression MLP with PyTorch

# installation
#install.packages("daltoolboxdp")

# loading DAL
library(daltoolbox)
library(daltoolboxdp)
```

Load Boston dataset (MASS) and inspect types.
```{r}
# Dataset for regression analysis

library(MASS)
data(Boston)
print(t(sapply(Boston, class)))
head(Boston)
```

Optional conversion to matrix.
```{r}
# for performance, you can convert to matrix
Boston <- as.matrix(Boston)
```

Random and reproducible train/test split.
```{r}
# preparing dataset for random sampling
set.seed(1)
sr <- sample_random()
sr <- train_test(sr, Boston)
boston_train <- sr$train
boston_test <- sr$test
```

Train MLP: define the hidden architecture and training controls.
```{r}
# Training

model <- torch_reg_mlp(
  attribute = "medv",
  input_size = ncol(Boston) - 1L,
  hidden_sizes = c(16L, 8L),
  epochs = 100L
)
model <- fit(model, boston_train)
```

Training evaluation.
```{r}
# Model adjustment

train_prediction <- predict(model, boston_train)
boston_train_predictand <- boston_train[, "medv"]
train_eval <- evaluate(model, boston_train_predictand, train_prediction)
print(train_eval$metrics)
```

Test evaluation.
```{r}
# Test

test_prediction <- predict(model, boston_test)
boston_test_predictand <- boston_test[, "medv"]
test_eval <- evaluate(model, boston_test_predictand, test_prediction)
print(test_eval$metrics)
```

Notes
- Default configuration uses `validation_strategy = "static"` and `stopping_rule = "none"`.
- To activate early stopping, set `stopping_rule` to `"patience"`, `"sma"`, `"ema"`, or `"h"`.
- To activate dynamic validation splits, use `validation_strategy = "dynamic"`.

References
- Bishop, C. M. (1995). Neural Networks for Pattern Recognition. Oxford University Press.
- Paszke, A., et al. (2019). PyTorch: An Imperative Style, High-Performance Deep Learning Library.
